Envelope Estimation using Geometric Properties of a Discrete Real Signal

### 2021-05-10 Papers/01_Pulses_Envelope/00ElsevierPulsesEnvelope.tex
revising

\maketitle
\linenumbers
\section{Introduction}
Envelope detection, also known as demodulation, is ubiquitous in both analogue and digital signal processing (DSP) \cite{2011CaetanoImproved}. Nevertheless, the literature in this area is fragmented \cite{2017LyonsDigital}, with most envelope detection techniques being designed to account for very specific settings, like pure sinusoids with moderate noise content, for example.

@@

In many contexts, however, the temporal amplitude envelope of a signal plays a prominent role: according to \cite{2017QiRelative}, for example, the envelope is at least as important as the fine structure of a sound wave in the context of the intelligibility of Mandarin tones. This is also the case for the English language \cite{1995ShannonSpeech}, where even envelopes modulating mostly noise are still capable of conveying meaning.

The temporal envelope also helps to impart emotion and identity to the human voice \cite{2018ZhuContributions} and music: Concert halls with envelope preserving characteristics are deemed more pleasant \cite{2011LokkiEngaging}, for example.

Specially when dealing with broadband signals, approaches tailored to specific applications, exploiting a priori knowledge of the waves, are prevalent, such as the one presented by \cite{2014YangFast} for the distributed monitoring of fibre optic or the one formulated by \cite{2018AssefModeling} in the context of medical ultrasound imaging.

@@

Considering the standard 2d representation of a discrete wave, where the horizontal axis represents time and the vertical axis represents the amplitude, one can think of an individual sample as a point. Provided a coherent mapping from this space to the standard Cartesian plane is available, the original DSP problem can be transformed into the geometric problem of finding the shape of a set of points.

One direct way of defining the shape of a set of points in $ \mathbb{R}^2 $ is using the concept of a convex hull, as defined in \cite{2008BergComputational}. The convex hull of a set of points $ S $ in two-dimensional space is the smallest convex subset of the plane that contains all the points in $ S $. Convex hulls are unique, in the sense that a given set of points define one and only one convex hull.

When the general shape of a set of points present prominent concavities, the convex hull overestimates the boundaries. To address this limitation \cite{1983Edelsbrunnershape} introduced the concept of alpha shapes, also known as concave hull: a mathematically well-defined generalization of the convex hull of a finite set of points, closely related to the Delaunay triangulation and Voronoi diagrams of those points.

% TODO
The intuition behind the alpha shapes algorithm is that only points that can be touched by a disc of radius $ \alpha $ coming from an infinite distance towards the baricenter of the set of points are part of the frontier. The polygon obtained linking those points forms the concave hull of this set.

Once a radius $ \alpha $ is defined, alpha shapes exhibit the uniqueness property: for a given set of points and a given radius, only one concave hull exists. When the radius $ \alpha $ of the circle tends to infinity, concave hulls tend to the convex hulls. This

Consider, for example, the artificially generated discrete wave in Fig. \ref{fig:Hulls} below, where the links between consecutive samples were deemphasized to highlight the “set of points” nature of a signal: two alpha shapes with different values for $ \alpha $ and the convex hull are illustrated. As $ \alpha $ increases, the contour of the set of points becomes smoother, and closer to the convex hull.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/00ConvexHull.pdf}
  \caption{A discrete signal interpreted as a set of points in the Cartesian plane, and the convex and concave hull at $ \alpha=300 $ defined by it.}
  \label{fig:Hulls}
\end{figure}

The alpha shapes approach is used in areas such as detection of features in images \cite{2016VarytimidisAlpha}, reconstruction of surfaces from a cloud of points \cite{2015WuAutomated} and spectroscopy \cite{2019XuModeling}.

This last work, which involves the estimation and removal of the Blaze function (a kind of envelope) of an echelle spectrograph, being particularly illustrative of the potential synergy between geometric and DSP approaches.

@@

In Section \ref{sec:representation} we explore some simplifications to the task of detecting the envelope of a signal that arise from the envelope definition adopted in this work and simplify the task of envelope detection.

Section \ref{sec:mapping} explores the geometric characteristics of envelope detection and formalizes how this classic DSP problem can be translated to the geometric problem of finding the shape of a set of points.

@@

Section \ref{sec:quality} discusses the difficulties in the assessment of the quality of an envelope, proposing qualitative and quantitative approaches, and demonstrates that the method propose conforms to four requirements for a physically plausible envelope proposed in \cite{1996Loughlinamplitude}.

In Section \ref{sec:comparison} we compare the presented algorithm with traditional envelope detection methods, both in terms of accuracy and time and present, in Section \ref{sec:extensions},  potential applications of the algorithm, some of which might not be immediately obvious.

% TODO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Interactive visualizations are available from the repository prepared for this work \cite{2020TarjanoEnvelope}, that also contains an implementation of the method in pure Python.
A Python module available at the Python Package Index (PyPI) installable via the command `pip install signal-envelope` for convenience, is also available. Besides, those working under Windows 64-bit can benefit from a specialized dynamic library coded in C++ that is automatically used within this module, and can be separately obtained via the release section in the repository.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Temporal envelope}\label{sec:envelope}

@@

While there is no general formal definition of the envelope of broadband signals \cite{2014Yangnovel,2015Yangtheoretical,2019Jiaempirical}, the general intuition is that an envelope is a smooth curve situated as close as possible to the signal \cite{2019Jiaempirical}.

@@

Fig. \ref{fig:AnalyticSignal} exemplifies the Hilbert envelope, for a pure sinusoid modulated by a polynomial, illustrating the difference in the smoothness of the envelope in the presence and absence of Gaussian white noise. It is readily noticeable from the Fig. \ref{fig:AnalyticSignal} that, as soon as noise is introduced in the original signal, it is directly reflected in the envelope.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/01AnalyticSignal.pdf}
  \caption{Envelope of a pure sinusoid with a local frequency of 20 cycles modulated by a polynomial of degree 3, as obtained by the Hilbert Transform approach. In the first half, the sinusoid is free of noise, while in the second half Gaussian noise with a standard deviation of $ 1/10 $ of the maximum amplitude of the wave was added to the base signal.}
  \label{fig:AnalyticSignal}
\end{figure}

For a real world signal, such as an alto singer sustaining a steady note as shown in Fig. \ref{fig:AnalyticSignalAlto} the Hilbert transform alone provides a poor envelope, since it retains most of the frequencies of the underlying wave. With further processing, however, such as filtering, a smoother output can be obtained.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/01AnalyticSignalAlto.pdf}
  \caption{Representation of an alto singer sustaining a steady note and a comparison of the envelopes obtained by the Hilbert transform with and without further filtering.}
  \label{fig:AnalyticSignalAlto}
\end{figure}

@@

$ E(t) $ represents the slow varying part of the wave, also known as the (temporal) envelope, modulator component, or amplitude modulation (AM) of $ W(t) $, while $ C(t) $ models its fast varying part, called throughout the literature as its (temporal) fine structure, carrier wave, or frequency modulation (FM).

@@

We address this problem by assuming $ C(t) $ normalized between $ \{-1, 1\}$ or, in other words $ |\max(C(t))| = 1 $. This assumption doesn't cause any loss of generality since, given an arbitrary $ C(t) $, we can obtain its normalization by dividing the function by its absolute global maximum, that is, $ \hat{C}(t) = C(t) / \max(|C(t)|) $, provided that $ \max(|C(t)|) \ne 0 $.

In this work we are concerned with the discrete version of this problem: given a finite digital wave represented by the vector $ \mathbf{w} \in \mathbb{R}^n $, instead of the continuous function $ W(t) $, obtaining the temporal envelope $ \mathbf{e} $ of $ \mathbf{w} $.

@@

\begin{align} \label{eq:Envelope}
\begin{split}
  \text{Given:}& \\
  \mathbf{w} & \in \mathbb{R}^n, \\
  \mathbf{w} & = (w_0, w_1, \cdots, w_{n-1}) \\
  \text{We define:}& \\
  \mathbf{e}, \mathbf{c} & \in \mathbb{R}^n \\
  \mathbf{e} & = (e_0, e_1, \cdots, e_{n-1}), w_i \ge 0 \ \forall \ i, \ 0 \le i \le n-1 \\  
  \mathbf{c} & = (c_0, c_1, \cdots, c_{n-1}), |c_i| \le 1 \ \forall \ i, \ 0 \le i \le n-1 \\
  \mathbf{w} & = \mathbf{e} \odot \mathbf{c} \ \therefore \ w_i = e_i c_i \ \forall \ i, \ 0 \le i \le n-1
\end{split}
\end{align}

@@

In Fig. \ref{fig:Envelope} a signal arising from the multiplication of a known carrier wave \textbf{c} and a known envelope \textbf{e} is shown; this helps to illustrate a key concept: note from Fig. \ref{fig:Envelope} that, only at some local extrema, the envelope can be inferred without prior knowledge of the underlying carrier, being in fact equal to the value of the wave $ w_i $. $ c_i = 1 $ at those points, rendering $ e_i = 1 w_i $.

This realization that the envelope of a wave is defined only at some of the wave's local extrema is crucial, as we can use it to generate a simplified representation of a signal that will make the algorithm more efficient by considering only the local extrema of a given signal.

\section{Simplified Representation} \label{sec:representation}

We start with the definition of a pulse: each time the value of $ w_i$ changes sign, that is, every time the amplitude of the discrete wave \textbf{w} crosses the horizontal axis, the beginning of a pulse is defined, with the next crossing defining its end. This definition is in line with the one presented in \cite{national1996telecommunications}, where a pulse is defined as a rapid change in the amplitude of a signal, followed by a fast return to the baseline value; zero, in our case.

From \ref{fig:Envelope} and the discussion in Section \ref{sec:envelope}, we can see that only the maximum point of a positive pulse or the minimum point of a negative pulse can potentially be equal to the envelope and are, thus, our only points of interest. We can then proceed, for the rest of the method, considering only those points, to great computational economy.

@@

This relation is illustrated in Fig. \ref{fig:Pulses}. We call it P to emphasize the connection with the pulses of \textbf{w}.
While the exact relation between $ n $, the size of the original signal, and $ m $, the size of the set of points, is dependent on factors such as the frequency and sampling rate of the original wave, the inequality $ m \le n $ holds for all discrete signals. In practice, $ m \ll n $ for most real world signals.

@@

To achieve that, we divide the ordinate of each point by the average of all ordinates, effectively cancelling the unit of the vertical components of the points. If we proceeded similarly in relation to the horizontal components, 
the direct link between P and \textbf{w} that arises from the fact that the abscissas of the points in P are numerically equivalent to the indices $ i $ of \textbf{w} would be lost.

It is best to leave the abscissas intact and multiply the now unitless vertical coordinates of each point by the average of the difference between the abscissa of a point and the abscissa of the immediately posterior. In this way, we can consider both axes as having unit $ i $, and the relationship is preserved, making it easier to recover the envelope points in the original basis later.

@@

\begin{align}\label{eq:Basis}
\begin{split}
P_{\mathbf{b}} &= (x_1, x_2) \\
\beta &= \frac{\left( \frac{i_{m-1} - i_m}{m-1} \right)}{\left( \frac{\sum_{j=0}^{m-1} \lvert w_j \lvert}{m} \right)}  \\
P_{\mathbf{s}} &= (x_1, \beta x_2) \quad \forall \  P \in \text{P} \\ 
\end{split}
\end{align}

The effect of this normalization can be seen in the points shown in Fig. \ref{fig:Coordinates}, shown in scale for both basis. We can consider our points now residing in the Cartesian coordinate system.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/04Coordinates.pdf}
  \caption{The set of points P, in both the original (top) and standard (bottom) coordinate basis, shown in scale.}
  \label{fig:Coordinates}
\end{figure}

\section{Discrete Curvature Estimation} \label{sec:curvature}

Having established a transformation from the original space of the two-dimensional representation of a discrete wave to the Cartesian coordinate system and, thus, a way to consistently transform our simplified description of a discrete signal into a set of points in the plane, the shape of which can be estimated with the use of geometric algorithms.

The goal of those algorithms, such as the convex hull and the alpha-shapes, is to discriminate which points of a set are part of its frontier.

A formal definition of alpha-shapes is given at \cite{1983Edelsbrunnershape}. In the context of this work, gaining the intuition of the algorithm is more important, and thus the explanation of the alpha-shapes algorithm presented in \cite{1994EdelsbrunnerThree} can be more insightful.

The explanation, translated to our context, invites one to think of the plane were the samples are represented as a soft material, such as Styrofoam. The samples themselves are a harder material, such as rock, and occupy a fixed position in this plane. The alpha shapes algorithm consists of carving this plane from the outside of those points, with a circular eraser of fixed radius $ \alpha $. In the end of the process, the remaining Styrofoam that couldn't be erased due to the impeding rocks is the alpha-shape of this set of points.

Alternatively, one can picture a circle of fixed radius being rolled above the signal, and marking the points it touches as envelope points.

To employ an alpha shapes inspired algorithm in this set of points, however, the radius $ r $ of such circle must be estimated, and for that a measure of discrete curvature is needed.

@@

The two general approaches are the derivation of direct methods that use characteristics of the discrete curve to calculate the curvature, or the use of the curvature of a smooth, continuous curve fitted to the original discrete curve \cite{2001CoeurjollyDiscrete}.

Concerning direct methods \cite{2014CarrollSurvey}, for example, derive three such definitions based on the approximation of a circle by an inscribed, centered and circumscribed polygon. In the context of three-dimensional meshes, \cite{2016VasaMultivariate} evaluate a range of existing estimators from a multivariate point of view.

Those approaches define the discrete curvature in relation to the vertices of a discrete curve (or mesh, in the three-dimensional case). Instead, we are interested in the change of direction, what the term curvature ultimately means, between two adjacent points.

Since we are dealing with a potentially large set of points, it is important for this measure to act locally, to assure $ O(n) $ complexity.

@@

\begin{figure}[ht!]
  \centering
    \includegraphics{images/04Vectors.pdf}
  \caption{The set of points P and the set V of vectors between two adjacent points of P, in the normalized coordinate system.}
  \label{fig:Vectors}
\end{figure}

We formalize the set of vectors V in Eq. \ref{eq:Vectors} below.

>>>>

\maketitle
\linenumbers

\section{Introduction}
Envelope detection, also known as demodulation, is ubiquitous in both analog and digital signal processing (DSP) \cite{2011CaetanoImproved}. Nevertheless, the literature in this area is fragmented \cite{2017LyonsDigital}, with most envelope detection techniques being designed to account for very specific settings, in general involving narrowband signals with moderate noise content.

@@

In many contexts, however, the temporal amplitude envelope of a signal plays a prominent role: according to \cite{2017QiRelative}, for example, the envelope is at least as important as the fine structure of a sound wave in the context of intelligibility of Mandarin tones. This is also the case for the English language \cite{1995ShannonSpeech}, where even envelopes modulating mostly noise are still capable of conveying meaning.

The temporal envelope also helps to impart emotion and identity to the human voice \cite{2018ZhuContributions} and music: Concert halls with envelope preserving characteristics are deemed more pleasant \cite{2011LokkiEngaging}.

Specially when dealing with broadband signals, approaches tailored to specific applications, exploiting prior knowledge of the waves, are prevalent, such as the one presented by \cite{2014YangFast} for the distributed monitoring of fiber optic or the one formulated by \cite{2018AssefModeling} in the context of medical ultrasound imaging.

@@

Considering the standard graphic representation of a discrete wave, where the horizontal axis represents time and the vertical axis represents the amplitude, one can think of each sample as a point in a plane. Provided a coherent mapping from this space to the Cartesian plane is available, the original DSP problem can be transformed into the geometric problem of finding the shape of this set of points.

One direct way of defining the shape of a set of points in $ \mathbb{R}^2 $ is by using the concept of a convex hull, as defined in \cite{2008BergComputational}. The convex hull of a set of points $ S $ in two-dimensional space is the smallest convex subset of the plane that contains all the points in $ S $. Convex hulls are unique, in the sense that a given set of points define one and only one convex hull.

When the general shape of a set of points presents prominent concavities, the convex hull overestimates the boundaries. To address this limitation, \cite{1983Edelsbrunnershape} introduced the concept of alpha shapes, also known as concave hulls: a mathematically well-defined generalization of the convex hull of a finite set of points, closely related to the Delaunay triangulation and Voronoi diagrams of those points.

The intuition behind the alpha shapes algorithm is that, given a set of points and a fixed radius $ \alpha $, if it is possible to construct a circle containing one or more points of the set in such a way that the disc defined by the circle does not contain any other point of the set, then these points are part of the frontier. The polygon obtained linking all such points forms the alpha shape, or concave hull, of this set.

Fig. \ref{fig:AlphaShape} illustrates an alpha shape, for $ \alpha=10 $, of a set of points. Note that two points of the set lie at the circle illustrated, but the open disc defined by it contains no points of the set. An analogous construction is possible for every point at the frontier. For all the interior points, it is impossible to construct a circle of radius 10 that contains one or more of them without defining a disc containing at least one point of the set.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/00AlphaShape.pdf}
  \caption{The alpha shape of a set of points for $ \alpha=10 $. A circle of radius 10, containing two points of the set, but defining a disc that doesn't contain any point of the set, is also shown.}
  \label{fig:AlphaShape}
\end{figure}

Once a radius $ \alpha $ is defined, alpha shapes exhibit the uniqueness property: for a given set of points and a given radius, only one alpha shape exists. When the radius $ \alpha $ of the circle tends to infinity, alpha shapes, also known as concave hulls, tend to the convex hulls.

Consider, for example, the artificially generated discrete wave in Fig. \ref{fig:Hulls} below, where the links between consecutive samples were de-emphasized to highlight the “set of points” nature of a signal: two alpha shapes, with different values for $ \alpha $, and the convex hull are illustrated. As $ \alpha $ increases, the contour of the set of points becomes smoother, and closer to the convex hull.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/00ConvexHull.pdf}
  \caption{A discrete signal interpreted as a set of points in the Cartesian plane, two alpha shapes, and the convex hull defined by it.}
  \label{fig:Hulls}
\end{figure}

The alpha shapes theory is used in areas such as detection of features in images \cite{2016VarytimidisAlpha}, reconstruction of surfaces from a cloud of points \cite{2015WuAutomated} and spectroscopy \cite{2019XuModeling}.

This last work, which involves the estimation and removal of the Blaze function (a kind of envelope) of an echelle spectrograph, is particularly illustrative of the potential synergy between geometric and DSP approaches.

@@

In Section \ref{sec:representation} we explore some simplifications to the task of detecting the envelope of a signal that arise from the envelope definition adopted in this work.

Section \ref{sec:mapping} explores the geometric characteristics of envelope detection and formalizes how this classic DSP problem can be transformed into the geometric problem of finding the shape of a set of points.

@@

Section \ref{sec:quality} discusses the difficulties in the assessment of the quality of an envelope, proposing qualitative and quantitative approaches, and demonstrates that the method presented conforms to four requirements for a physically plausible envelope proposed in \cite{1996Loughlinamplitude}.

In Section \ref{sec:comparison} we compare the presented algorithm with traditional envelope detection methods in terms of accuracy and time and present, in Section \ref{sec:extensions},  potential applications of the algorithm, some of which might not be immediately obvious.

Links to interactive visualizations are available from the repository prepared for this work \cite{2020TarjanoEnvelope}, which also contains an implementation of the method in pure Python.
A Python module can be installed from the Python Package Index (PyPI) with the command `pip install signal-envelope`. Besides, those working under Windows 64-bit can benefit from a specialized dynamic library coded in C++ that is automatically used within this module, and can be separately obtained via the release section in the repository.

\section{Temporal envelopes}\label{sec:envelope}

@@

While there is no general formal definition for the envelope of broadband signals \cite{2014Yangnovel,2015Yangtheoretical,2019Jiaempirical}, the consensus is that an envelope is a smooth curve situated as close as possible to the signal \cite{2019Jiaempirical}.

@@

Fig. \ref{fig:AnalyticSignal} exemplifies the Hilbert envelope, for a pure sinusoid modulated by a polynomial, illustrating the difference in the smoothness of the envelope in the presence and absence of Gaussian white noise. It is readily noticeable from Fig. \ref{fig:AnalyticSignal} that, as soon as noise is introduced in the original signal, it is directly reflected in the envelope.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/01AnalyticSignal.pdf}
  \caption{Envelope of a pure sinusoid with a local frequency of 20 cycles, modulated by a polynomial of degree 3, as obtained by the Hilbert Transform approach. In the first half of the samples, the sinusoid is free of noise, while in the second half Gaussian noise, with a standard deviation of $ 1/10 $ of the maximum amplitude of the wave, was added to the base signal.}
  \label{fig:AnalyticSignal}
\end{figure}

For a real world signal, such as an alto singer sustaining a steady note as shown in Fig. \ref{fig:AnalyticSignalAlto}, the Hilbert transform alone provides a poor envelope, since it retains most of the frequencies of the underlying wave. With further processing, however, such as filtering, a smoother output can be obtained.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/01AnalyticSignalAlto.pdf}
  \caption{A section of the representation of an alto singer sustaining a steady note and a comparison of the envelopes obtained by the Hilbert transform with and without further filtering.}
  \label{fig:AnalyticSignalAlto}
\end{figure}

@@

$ E(t) $ represents the slow varying part of the wave, also known as the (temporal) envelope, modulator component, or amplitude modulation (AM) of $ W(t) $, while $ C(t) $ models its fast-varying part, called throughout the literature as its (temporal) fine structure, carrier wave, or frequency modulation (FM).

@@

We address this problem by assuming $ C(t) $ to be normalized between $ \{-1, 1\}$ or, in other words $ |\max(C(t))| = 1 $. This assumption does not cause any loss of generality since, given an arbitrary $ C(t) $, we can obtain its normalization by dividing the function by its absolute global maximum, that is, $ \hat{C}(t) = C(t) / \max(|C(t)|) $, provided that $ \max(|C(t)|) \ne 0 $.

In this work we are concerned with the discrete version of the envelope detection problem: given a finite digital wave represented by the vector $ \mathbf{w} \in \mathbb{R}^n $, obtaining the temporal envelope $ \mathbf{e} $ of $ \mathbf{w} $.

@@

\begin{align} \label{eq:Envelope}
\begin{split}
  \text{Given:}& \\
  \mathbf{w} & \in \mathbb{R}^n, \\
  \mathbf{w} & = (w_0, w_1, \cdots, w_{n-1}) \\
  \text{We define:}& \\
  \mathbf{e}, \mathbf{c} & \in \mathbb{R}^n \\
  \mathbf{e} & = (e_0, e_1, \cdots, e_{n-1}), e_i \ge 0 \ \forall \ i, \ 0 \le i \le n-1 \\  
  \mathbf{c} & = (c_0, c_1, \cdots, c_{n-1}), |c_i| \le 1 \ \forall \ i, \ 0 \le i \le n-1, |\max(\mathbf{c})| =1 \\
  \mathbf{w} & = \mathbf{e} \odot \mathbf{c} \ \therefore \ w_i = e_i c_i \ \forall \ i, \ 0 \le i \le n-1
\end{split}
\end{align}

@@

In Fig. \ref{fig:Envelope} a signal arising from the multiplication of a known carrier wave \textbf{c} and a known envelope \textbf{e} is shown; this helps to illustrate a key concept: note from Fig. \ref{fig:Envelope} that, only at some local extrema, the envelope can be inferred without prior knowledge of the underlying carrier, being, in fact, equal to the value of the instantaneous amplitude of the wave $ w_i $. That is, $ c_i = 1 $ at those points, rendering $ e_i = 1 w_i $.

This realization that the envelope of a wave is defined only at some of the wave's local extrema is important, as we can use it to generate a simplified representation of a signal that will make the algorithm more efficient.

\section{Simplified Representation of a signal} \label{sec:representation}

We start with the definition of a pulse: each time the value of $ w_i $ changes sign, that is, every time the instantaneous amplitude of the discrete wave \textbf{w} crosses the horizontal axis, the beginning of a pulse is defined, with the next crossing defining its end. This interpretation of a pulse is in line with the one presented in \cite{national1996telecommunications}, where a pulse is defined as a rapid change in the amplitude of a signal, followed by a fast return to the baseline value; zero, in our case.

From Fig. \ref{fig:Envelope} and the discussion in Section \ref{sec:envelope}, we can see that only the maximum point of a positive pulse or the minimum point of a negative pulse can potentially be equal to the envelope and are, thus, our only points of interest. We can then proceed, for the rest of the method, considering only those points, to a great computational economy.

@@

This relation is illustrated in Fig. \ref{fig:Pulses}. We call it P to emphasize the connection with the pulses of \textbf{w}.
While the exact relation between $ n $, the number of samples of the original signal, and $ m $, the cardinality of the set of points, is dependent on factors such as the frequency and sampling rate of the original wave, the inequality $ m \le n $ holds for all discrete signals. In practice, $ m \ll n $ for most real-world signals.

@@

To achieve that, we divide the ordinate of each point by the average of all ordinates, effectively canceling the unit of the vertical components of the points. If we proceeded similarly in relation to the horizontal components,
the direct link between P and \textbf{w} that arises from the fact that the abscissas of the points in P are numerically equivalent to the indices $ i $ of \textbf{w} would be lost.

It is best to leave the abscissas intact and multiply the now unitless vertical coordinates of each point by the average of the difference between the abscissa of a point and the abscissa of the immediately posterior. In this way, we can consider both axes as having unit $ i $, and the relationship is preserved, making it easier to recover the envelope points in the original coordinates later.

@@

\begin{align}\label{eq:Basis}
\begin{split}
P_{\mathbf{b}} &= (x_1, x_2) \\
\beta &= \frac{\left( \frac{i_{j-1} - i_j}{m-1} \right)}{\left( \frac{\sum_{j=0}^{m-1} \lvert w_j \lvert}{m} \right)}  \\
P_{\mathbf{s}} &= (x_1, \beta x_2) \quad \forall \  P \in \text{P} \\ 
\end{split}
\end{align}

The effect of this transformation can be seen in the points represented in Fig. \ref{fig:Coordinates}, shown in scale for both coordinate systems. We can consider our points residing in the Cartesian coordinate system after the transformation.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/04Coordinates.pdf}
  \caption{The set P of points, in both the original (top) and standard (bottom) coordinate systems, shown in scale.}
  \label{fig:Coordinates}
\end{figure}

\section{Discrete curvature estimation} \label{sec:curvature}

Having established a transformation from the original space of the two-dimensional representation of a discrete wave to the Cartesian coordinate system, we are now in a position to apply the geometric concept of the shape of a set of points.

Algorithms such as the convex hull and the alpha shapes solve that problem by discriminating which points of a set are part of its frontier.

A formal definition of alpha shapes is given at \cite{1983Edelsbrunnershape}. In the context of this work, however, gaining the intuition of the algorithm is more important, and thus the explanation presented in \cite{1994EdelsbrunnerThree} can be more insightful.

The explanation, translated to our context, invites one to think of the plane where the points of P are represented as a soft material, such as Styrofoam. The points themselves are made of a harder material, such as rock, and occupy a fixed position in this plane. The alpha shapes algorithm consists of carving this plane from the outside of those points, with a circular eraser of fixed radius $ \alpha $. At the end of the process, the remaining Styrofoam that couldn't be erased due to the impeding rocks is the alpha-shape of this set of points.

Alternatively, one can picture a circle of fixed radius being rolled above the set P, and marking the points it touches as part of the envelope.

To employ an alpha shapes inspired algorithm, however, the radius $ r $ of such a circle must be estimated, and for that, a measure of discrete curvature is needed.

@@

The two general approaches are the derivation of direct methods that use characteristics of the discrete curve to calculate the curvature, or the use of the curvature of a smooth, continuous curve fitted to the original discrete curve \cite{2001CoeurjollyDiscrete}. Since we are dealing with a potentially large set of points, it is important for this measure to act locally, to assure $ O(n) $ complexity, as is generally not the case for approaches that involve curve fitting.

Concerning direct methods \cite{2014CarrollSurvey}, for example, derive three such definitions based on the approximation of a circle by an inscribed, centered, and circumscribed polygon. In the context of three-dimensional meshes, \cite{2016VasaMultivariate} evaluate a range of existing estimators from a multivariate point of view.
Those approaches define the discrete curvature in relation to the vertices of a discrete curve (or mesh, in the three-dimensional case). Since we are interested in the change of direction, what the term curvature ultimately means, between two adjacent points, a definition of discrete curvature over the edges of a discrete curve would be preferable.

@@

\begin{figure}[ht!]
  \centering
    \includegraphics{images/04Vectors.pdf}
  \caption{The set P of points and the set V of vectors between two adjacent points of P, in the Cartesian coordinate system.}
  \label{fig:Vectors}
\end{figure}

We formalize the set V of vectors in Eq. \ref{eq:Vectors} below.
